---
title: "Assignment3"
author: "RF"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

The questions you will have to answer are:

1. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results. P.S. Should you scale? P.P.S. Don't forget to motivate your priors. P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.
1.1. Verbal IQ and ADOS
1.2. Non Verbal IQ and ADOS
1.3. Social IQ and ADOS
```{r}
#Load data
data = read.csv("Assignment3Data.csv")
library(rethinking)

#we need to make a model-stuffing-thingy
#I will make a data subset only for visit 1
dataSub = data[data$Visit == 1,]

#We remove participant 1, with NAs
dataSub = dataSub[dataSub$ChildID != 1,]

#Now only children with ASD
dataSub = dataSub[dataSub$ASD == 1,]


#----------VERBALIQ----------#

#Check the distribution
plot(density(dataSub$VerbalIQ))

#Symptom severity = ADOS = our model outcome
#IQ = verbal, nonverbal and social IQ -> We predict all to be correlated with the data
#So we have:
  # ADOS_i ~ Normal(mu_i, sigma)
  # mu = a + b*verbalIQ
#Our priors
  #alpha ~ dnorm(20,10) - this is the mean of ados when x = 0 (verbalIQ)
  #Beta ~ normal(0,10) -> Stregnth of the relation between ADOS and verbal IQ -> No clue why we would choose these values #zero because we do not know what we expect
  #sigma ~ uniform(0, 30) #30 er circa max

mVIQ <- map(
  alist(
    ADOS ~ dnorm(mu, sigma),
    mu <- a + b*VerbalIQ,
    a ~ dnorm(20, 10),
    b ~ dnorm(0, 10),
    sigma ~ dunif(0,30)
  ), data = dataSub
)
precis(mVIQ)

plot(ADOS ~ VerbalIQ, data = dataSub, col = rangi2)
abline(a = coef(mVIQ)["a"], b= coef(mVIQ)["b"])

#A plot with uncertainty
post <- extract.samples(mVIQ, n = 20)

#Display raw data and sample size
plot(dataSub$VerbalIQ, dataSub$ADOS,
     xlim = range(dataSub$VerbalIQ), ylim = range(dataSub$ADOS),
     col = rangi2, xlab = "Verbal IQ", ylab = "ADOS")
#plot the lines
for (i in 1:20)
  abline(a = post$a[i], b = post$b[i], col = col.alpha("black", 0.3))



#Plot our data against the distribution
#First simulate
sim.ADOS.vIQ <- sim(mVIQ, data = dataSub, n = 1e4)

#dens plot on raw data
dens(sim.ADOS.vIQ, col = "red", xlim = c(0, 30), ylim = c(0,0.12),  xlab = "ADOS")
par(new=TRUE)
dens(dataSub$ADOS, xlim = c(0, 30), ylim = c(0,0.12), xlab = "ADOS")
title("Verbal IQ")
  

```


2. Do the different aspects of IQ account for different portions of the variance in ADOS? 
2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?
2.2. Build the model, assess its quality, write a few lines interpreting the results.

```{r}
#First we want to be able to do the precis stuff
library(rethinking)


#So now we standardize
dataSub$VerbalIQ.s <- scale(dataSub$VerbalIQ)[,]
dataSub$NonVerbalIQ.s <- scale(dataSub$NonVerbalIQ)[,]
dataSub$SocialIQ.s <- scale(dataSub$SocialIQ)[,]
dataSub$ADOS.s <- scale(dataSub$ADOS)[,]


#Do the MAP-thingy
m2.1 <- map(
  alist(
    ADOS.s ~ dnorm(mu, sigma),
    mu <- a + bV*VerbalIQ.s + bN*NonVerbalIQ.s+bS*SocialIQ.s,
    a ~ dnorm(0, 2),
    bV ~ dnorm(0,1),
    bN ~ dnorm(0, 1),
    bS ~ dnorm(0,1),
    sigma ~ dunif(0,10)), data = dataSub
)

precis(m2.1, corr = T)
plot(precis(m2.1))

#by centering we get away from the correlations
#Center everything
#Standardizing puts everything on the same scale.
#we can choose to do both
#non-verbal IQ overlaps with zero - not necessarily any added information from adding this (As it overlaps with zero)

pairs( ~ ADOS.s + VerbalIQ.s + NonVerbalIQ.s + SocialIQ.s, data = dataSub, col = rangi2)

#Different ways:
#- testing
# conceptual
#Crying
# FIrst answer:IQ might be the expression for an underlying G-factor -> They might be colinear. THeory driven hypothesis. If they are fully correlated, we would expect the plots of the coefficients to be centered around zero
#We might have an alternative hypothesis: Different intelligences that might not be corralted -> Prediction: adding a new IQ-type will always add more value

#v ~ nv -> Find the beta, beta = .67
#v ~s -> beta = .6
# Nv ~ s  beta = .37

#Three aspects. strong connection between verbal and nonverbal, weaker connection between social and nonverbal iQ


#Plot our data against the distribution
#First simulate
sim.ADOS <- sim(m2.1, data = dataSub, n = 1e4)

#dens plot on raw data
dens(sim.ADOS, col = "red", xlim = c(-3, 3), ylim = c(0,0.55),  xlab = "ADOS")
par(new=TRUE)
dens(dataSub$ADOS.s, xlim = c(-3, 3), ylim = c(0,0.55), xlab = "ADOS")
title("Plot")






#Look how much the posterior beta moves of each predictor from being alone (question 1) when we then have the others


m2.2 <- map(
  alist(
    VerbalIQ.s ~ dnorm(mu, sigma),
    mu <- a + bN*NonVerbalIQ.s,
    a ~ dnorm(0, 10),
    bN ~ dnorm(0, 1),
    sigma ~ dunif(0,10)), data = dataSub
)

precis(m2.2, corr = T)
plot(precis(m2.2))

#Calculate the error: Difference between the two distribution. Or subtract one distribution from the other
#look at whcih one has the biggest sigma (little bit cheating, does not take into account how precise the beta-values are.)


```


3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.

4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?
4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.

5. Bonus questions: Including measurement errors. 
5.1. Let's assume that ADOS has a measurement error of 1. How would you model that in any of the previous models? 
5.2. We know that IQ has an estimated measurement error of 2.12. How would you include that? 


